### 2020-04-17 tidymodels/learning_tidymodels.Rmd
tutorial comparativo Caret vs Tidymodels

---
title: "Learning Tidymodels"
output: html_notebook
---

Esse notebook R tem como finalidade conhecer e exercitar o conjunto de pacotes `tidymodel`, que seria uma evolução do pacote `caret` para transformação e fitting de dados.

Reproduz ou se baseia nos seguintes tutoriais da internet:

* [Caret vs. tidymodels - comparing the old and new](https://konradsemsch.netlify.app/2019/08/caret-vs-tidymodels-comparing-the-old-and-new/)
* [IMPLEMENTING THE SUPER LEARNER WITH TIDYMODELS](https://www.alexpghayes.com/blog/implementing-the-super-learner-with-tidymodels/)

# Seguindo o exemplo

## Bibliotecas

```{r loadlib, message=FALSE, warning=FALSE}
set.seed(42)
options(max.print = 150)

library(modeldata)
library(tidymodels)
library(tidyverse)
library(caret)
library(magrittr)
library(naniar)
library(furrr)
library(skimr)
library(vip)
library(workflows)
library(tune)

plan(multicore)  
```

## Data preparation

```{r}

data("credit_data")

credit_data <- credit_data %>% set_names(tolower(names(.)))

glimpse(credit_data)

```

## Checking missing data com `naniar` e avalinado

```{r}

credit_data %>% 
  miss_var_summary()

```

## avaliando os dados

```{r}

credit_data %>% skim()

```

## Class Balance

```{r}
table(credit_data$status)
round(prop.table(table(credit_data$status)),2)
```

## modeling

```{r}

split <- rsample::initial_split(credit_data, prop = .8, strata = "status")

df_train <- training(split)
df_test  <- testing(split)

train_cv <- rsample::vfold_cv(df_train, v=5, strata = "status")
train_cv_caret <- rsample2caret(train_cv)

```

In this particular example I mainly focus on imputting missing data or assigning them a new categorical level, infrequent/ unobserved values and hot-encoding them.

```{r}

my_recipe <- df_train %>% 
  recipe(status~.) %>% 
  # imputation: add "unknown" to all missing factor values
  step_unknown(all_nominal(), -status) %>% 
  # imputation: add median to all missing numeric values
  step_medianimpute(all_numeric()) %>% 
  # compining: group factors below 5% of frequency (default) in an "infrequent_combined"
  step_other(all_nominal(), -status, other = "infrequent_combined") %>% 
  # create a "level" in the factor columns for unseen factors
  step_novel(all_nominal(), -status, new_level = "unrecorded_observation") %>% 
  # OHE
  step_dummy(all_nominal(), -status, one_hot=T)

my_recipe
my_recipe_prep <- prep(my_recipe, retain=T)

my_recipe_prep
tidy(my_recipe_prep)

```

## playing with recipe

```{r}
# some data
mydata <- tibble(
  class = as.factor(c("dog","dog","dog","cat","cat","monkey")),
  size  = c(53,42,83,20,30,70),
  weight = c(15,8,12,5,6,21)
)

# create a recipe from "mydata"
rcp <- recipe(mydata) %>% 
  # one hot encoding fo class
  step_dummy(all_nominal(), one_hot = T)

# prepare the recipe
p_rcp <- prep(rcp, verbose = T)

# apply a prepared recipe to a new data
# in this case do OHE
bake(p_rcp, mydata)

```

## Caret

```{r}

control_caret <- trainControl(
  method="cv", 
  verboseIter = F,
  classProbs = T, 
  summaryFunction = twoClassSummary,
  returnResamp = "final",
  savePredictions = "final",
  index = train_cv_caret$index,
  indexOut = train_cv_caret$indexOut
)

grid_caret <- expand.grid(
  mtry = seq(1,ncol(df_train)-1,3),
  splitrule = c("extratrees","gini"),
  min.node.size=c(1,3,5)
)

model_caret <- train(
    status~.,
    data=juice(my_recipe_prep), 
    method="ranger", 
    metric="ROC",
    trControl=control_caret, 
    tuneGrid = grid_caret,
    importance="impurity",
    num.tree=500
  )

print(model_caret)
```

```{r}

varImp(model_caret, scale=T)$importance %>% 
  rownames_to_column() %>% 
  arrange(desc(Overall))

```

```{r}

df_train_pred_caret <- model_caret$pred %>% 
  group_by(rowIndex) %>% 
  summarise(bad=mean(bad)) %>% 
  transmute(estimate=bad) %>% 
  add_column(truth=df_train$status)

# cross-validated training performance
percent(roc_auc(df_train_pred_caret, truth, estimate)$.estimate)

```
```{r}
# test performance

df_test_pred_caret <- predict(
    model_caret, 
    newdata = bake(my_recipe_prep, df_test),
    type="prob"
  ) %>% 
  as_tibble() %>% 
  transmute(estimate=bad) %>% 
  add_column(truth=df_test$status)

percent(roc_auc(df_test_pred_caret, truth, estimate)$.estimate)

```

# tidymodel

```{r}
engine_tidym <- rand_forest(
    mode = "classification",
    mtry = tune(),
    trees = tune(),
    min_n = tune()
  ) %>% 
  set_engine("ranger",importance="impurity")

print(engine_tidym)
```

```{r}
gridtm <- grid_random(
  mtry() %>% range_set(c(1, 20)),
  trees() %>% range_set(c(500, 1000)), 
  min_n() %>% range_set(c(2, 10)),
  size = 30
  )
```

```{r}

wkfl_tidym <- workflow() %>% 
  add_recipe(my_recipe) %>% 
  add_model(engine_tidym)

gridy_tidym <- tune_grid(wkfl_tidym, resamples=train_cv, grid=gridtm,
                         metric=metric_set(roc_auc),
                         control=control_grid(save_pred=T))

```

```{r}
collect_metrics(gridy_tidym)
```

```{r}

grid_tidym_best <- select_best(gridy_tidym, metric="roc_auc")
(wkfl_tidym_best <- finalize_workflow(wkfl_tidym, grid_tidym_best))
```

```{r}
(wkfl_tidym_final <- last_fit(wkfl_tidym_best, split = split))
```

```{r}
# Cross-validated training performance
percent(show_best(gridy_tidym, metric="roc_auc", n = 1)$mean)
```

```{r}
# Test performance
percent(wkfl_tidym_final$.metrics[[1]]$.estimate[[2]])
```

```{r}
vip(pull_workflow_fit(wkfl_tidym_final$.workflow[[1]]))$data
```