### 2020-04-17 tidymodels/tutorial_in_tidymodels.Rmd
tutorial de Tidymodels

---
title: "Tutorial on Tidymodel"
output: md_document
---

Reproduzindo este [artigo](https://hansjoerg.me/2020/02/09/tidymodels-for-machine-learning/) sobre `tidymodels` do [Hansjörg Plieninger](https://hansjoerg.me/page/about/)

# Tutorial on tidymodels for Machine Learning

## Setup

```{r loadlibs, message=FALSE, warning=FALSE}
library(conflicted)
library(tidymodels)
library(tidyverse)
library(ggrepel)
library(corrplot)
library(skimr)

conflict_prefer("filter","dplyr")

ggplot2::theme_set(theme_light())
```

## Dataset: `Diamonds`

```{r diamond_plot}

data("diamonds")

head(diamonds,10)

diamonds %>% 
  sample_n(2000) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  select(price, everything()) %>% 
  cor() %>% 
  {.[order(abs(.[,1]), decreasing = T),
     order(abs(.[,1]), decreasing = T)]} %>% 
  corrplot(method = "number", type="upper", mar=c(0,0,1.5,0), 
           title="Correlations between price and various features of diamonds")

```

```{r}

diamonds %>% 
  skim()

```

## Separating Testing and Training Data: `rsample`

```{r}
# you know: the life, the universer... 
set.seed(42)

# training/test
dia_split <- initial_split(diamonds, prop = .1, strata = price)
dia_train <- training(dia_split)
dia_test  <- testing(dia_split)

dim(dia_train)
dim(dia_test)

# cv-folds
dia_vfold <- vfold_cv(dia_train, v=3, repeats=1, strata=price)

dia_vfold %>% 
  mutate(df_ana=map(splits, analysis),
         df_ass=map(splits, assessment)) -> folds

```

## Data Pre-Processing and Feature Engineering: `recipes`

The plot below indicates that there may be a nonlinear relationship between price and carat, and I want to address that using higher-order terms.

```{r}
qplot(carat, price, data = dia_train) +
    scale_y_continuous(trans = log_trans(), labels = function(x) round(x, -2)) +
    geom_smooth(method = "lm", formula = "y ~ poly(x, 4)") +
    labs(title = "Nonlinear relationship between price and carat of diamonds",
         subtitle = "The degree of the polynomial is a potential tuning parameter")
```

```{r}
dia_rec <- recipe(price~., data=dia_train) %>% 
  # log no preço
  step_log(all_outcomes()) %>% 
  # normializa valores (que fatoriais)
  step_normalize(all_predictors(), -all_nominal()) %>% 
  # OHE dos fatoriais
  step_dummy(all_nominal()) %>% 
  # cria uma feature (?) de polinomial
  step_poly(carat, degree = 4)

prep(dia_rec)

```

```{r}

# apply the recipe in the train data
dia_prep <- prep(dia_rec)

# extract transformed training data
dia_juiced <- juice(dia_prep)

# apply the recipe in a new data
bake(dia_prep, dia_test)

```

## Defining and Fitting Models: `parsnip`

The parsnip package has wrappers around many1 popular machine learning algorithms, and you can fit them using a unified interface. This is extremely helpful, since you have to remember only one rather then dozens of interfaces.

```{r}

# regression example
lm_model <- linear_reg() %>% 
  set_mode("regression") %>% 
  set_engine("lm")

rf_model <- rand_forest(mtry=3, trees=500, min_n=5) %>% 
  set_mode("regression") %>% 
  set_engine("ranger", importance="impurity")

```

```{r}

lm_fit <- fit(lm_model, price~., dia_juiced)
lm_fit

rf_fit <- fit(rf_model, price~., dia_juiced)
rf_fit
```

## Summarizing Fitted Models: broom

```{r}

glance(lm_fit$fit)
tidy(lm_fit$fit)
rf_fit$fit

augment(lm_fit$fit) %>% 
  rowid_to_column() %>% 
  select(rowid, price, .fitted:.std.resid)

```