### 2020-12-08 AdvendOfCode2020/draft_post.Rmd
Advent of Code 2020 Day 04 Part 2

---
title: "Advent Of Code 2020"
output: html_notebook
---

## Day 1: Report Repair

### Part One

After saving Christmas five years in a row, you've decided to take a vacation at a nice resort on a tropical island. Surely, Christmas will go on without you.

The tropical island has its own currency and is entirely cash-only. The gold coins used there have a little picture of a starfish; the locals just call them stars. None of the currency exchanges seem to have heard of them, but somehow, you'll need to find fifty of these coins by the time you arrive so you can pay the deposit on your room.

To save your vacation, you need to get all fifty stars by December 25th.

Collect stars by solving puzzles. Two puzzles will be made available on each day in the Advent calendar; the second puzzle is unlocked when you complete the first. Each puzzle grants one star. Good luck!

Before you leave, the Elves in accounting just need you to fix your expense report (your puzzle input); apparently, something isn't quite adding up.

Specifically, they need you to find the two entries that sum to 2020 and then multiply those two numbers together.

For example, suppose your expense report contained the following:

1721
979
366
299
675
1456

In this list, the two entries that sum to 2020 are 1721 and 299. Multiplying them together produces 1721 * 299 = 514579, so the correct answer is 514579.

Of course, your expense report is much larger. Find the two entries that sum to 2020; what do you get if you multiply them together?

```{r}
# solution using base r

# read the input as vector
input <- read.csv("day01_input.txt", header = F)[,1]

# generate all combination of 2 and generate a matrix
comb <- combn(input, 2)

# sum each combinations
sums <- colSums(comb)

# find which one has the sum equals 2020
vals <- comb[,sums==2020]

# multiply them
resp <- prod(vals)

resp

```

### Part two

The Elves in accounting are thankful for your help; one of them even offers you a starfish coin they had left over from a past vacation. They offer you a second one if you can find three numbers in your expense report that meet the same criteria.

Using the above example again, the three entries that sum to 2020 are 979, 366, and 675. Multiplying them together produces the answer, 241861950.

In your expense report, what is the product of the three entries that sum to 2020?

```{r}
# keeping within base r

# generic solution for any N combinations
findCombSum <- function(in.data, n.comb, match.value){

  # generate all combination of 'n.comb' and generate a matrix
  comb <- combn(in.data, n.comb)

  # sum each combinations
  sums <- colSums(comb)

  # find which one has the sum equals 'match.value'
  vals <- comb[,sums==match.value]

  # multiply them
  resp <- prod(vals)

  return(resp)

}

# read the input as vector
input <- read.csv("day01_input.txt", header = F)[,1]

# part 1
findCombSum(input, 2, 2020)

# part 2
findCombSum(input, 3, 2020)

```

## Day 2: Password Philosophy

### Parte One

Your flight departs in a few days from the coastal airport; the easiest way down to the coast from here is via toboggan.

The shopkeeper at the North Pole Toboggan Rental Shop is having a bad day. "Something's wrong with our computers; we can't log in!" You ask if you can take a look.

Their password database seems to be a little corrupted: some of the passwords wouldn't have been allowed by the Official Toboggan Corporate Policy that was in effect when they were chosen.

To try to debug the problem, they have created a list (your puzzle input) of passwords (according to the corrupted database) and the corporate policy when that password was set.

For example, suppose you have the following list:

1-3 a: abcde
1-3 b: cdefg
2-9 c: ccccccccc

Each line gives the password policy and then the password. The password policy indicates the lowest and highest number of times a given letter must appear for the password to be valid. For example, 1-3 a means that the password must contain a at least 1 time and at most 3 times.

In the above example, 2 passwords are valid. The middle password, cdefg, is not; it contains no instances of b, but needs at least 1. The first and third passwords are valid: they contain one a or nine c, both within the limits of their respective policies.

How many passwords are valid according to their policies?

```{r}
# tidyr::separate() and stringr::str_count() come in handy for this
library(tidyverse)

# read the input data as data frame with a column named 'input'
password.check <- read.csv("./day02_input.txt", header = F) %>% 
  set_names(c("input")) %>% 
  # split the line into the password and the policy columns
  separate(input, c("policy","password"), sep=": ", remove=F) %>% 
  # split the policy into range values and the letter to check
  separate(policy, c("pol.range.min", "pol.range.max","pol.letter"),
           sep="-| ", convert=T) %>% 
  # count the letter in the password and check the policy range
  mutate( count.letter = str_count(password, pol.letter),  
          is.valid = count.letter >= pol.range.min &
                     count.letter <= pol.range.max )

# let's see what we got
head(password.check)

# how many password are valid?
sum(password.check$is.valid)

```
### Part Two

While it appears you validated the passwords correctly, they don't seem to be what the Official Toboggan Corporate Authentication System is expecting.

The shopkeeper suddenly realizes that he just accidentally explained the password policy rules from his old job at the sled rental place down the street! The Official Toboggan Corporate Policy actually works a little differently.

Each policy actually describes two positions in the password, where 1 means the first character, 2 means the second character, and so on. (Be careful; Toboggan Corporate Policies have no concept of "index zero"!) Exactly one of these positions must contain the given letter. Other occurrences of the letter are irrelevant for the purposes of policy enforcement.

Given the same example list from above:

1-3 a: abcde is valid: position 1 contains a and position 3 does not.
1-3 b: cdefg is invalid: neither position 1 nor position 3 contains b.
2-9 c: ccccccccc is invalid: both position 2 and position 9 contain c.

How many passwords are valid according to the new interpretation of the policies?

```{r}
# tidyr::separate() and stringr::str_count() come in handy for this
library(tidyverse)

# read the input data as data frame with a column named 'input'
password.check <- read.csv("./day02_input.txt", header = F) %>% 
  set_names(c("input")) %>% 
  # split the line into the password and the policy columns
  separate(input, c("policy","password"), sep=": ", remove=F) %>% 
  # split the policy into range values and the letter to check
  separate(policy, c("pol.range.min", "pol.range.max","pol.letter"),
           sep="-| ", convert=T) %>% 
  # part 1 
  # count the letter in the password and check the policy range
  mutate( count.letter = str_count(password, pol.letter),  
          is.valid.part1 = count.letter >= pol.range.min &
                           count.letter <= pol.range.max ) %>% 
  # part 2
  # get the letters in positions 'min' and 'max"' and "XOR" check
  mutate(
    letter.at.pos1 = str_sub(password, pol.range.min, pol.range.min),
    letter.at.pos2 = str_sub(password, pol.range.max, pol.range.max),
    is.valid.part2 = xor(letter.at.pos1==pol.letter, letter.at.pos2==pol.letter)
  )

# part one: how many password are valid (repeat letter policy)?
sum(password.check$is.valid.part1)

# part tow: how many password are valid (letter at position xor policy?
sum(password.check$is.valid.part2)

```
## Day 3: Toboggan Trajectory

### Part 1

With the toboggan login problems resolved, you set off toward the airport. While travel by toboggan might be easy, it's certainly not safe: there's very minimal steering and the area is covered in trees. You'll need to see which angles will take you near the fewest trees.

Due to the local geology, trees in this area only grow on exact integer coordinates in a grid. You make a map (your puzzle input) of the open squares (.) and trees (#) you can see. For example:

```
..##.......
#...#...#..
.#....#..#.
..#.#...#.#
.#...##..#.
..#.##.....
.#.#.#....#
.#........#
#.##...#...
#...##....#
.#..#...#.#
```

These aren't the only trees, though; due to something you read about once involving arboreal genetics and biome stability, the same pattern repeats to the right many times:

```
..##.........##.........##.........##.........##.........##.......  --->
#...#...#..#...#...#..#...#...#..#...#...#..#...#...#..#...#...#..
.#....#..#..#....#..#..#....#..#..#....#..#..#....#..#..#....#..#.
..#.#...#.#..#.#...#.#..#.#...#.#..#.#...#.#..#.#...#.#..#.#...#.#
.#...##..#..#...##..#..#...##..#..#...##..#..#...##..#..#...##..#.
..#.##.......#.##.......#.##.......#.##.......#.##.......#.##.....  --->
.#.#.#....#.#.#.#....#.#.#.#....#.#.#.#....#.#.#.#....#.#.#.#....#
.#........#.#........#.#........#.#........#.#........#.#........#
#.##...#...#.##...#...#.##...#...#.##...#...#.##...#...#.##...#...
#...##....##...##....##...##....##...##....##...##....##...##....#
.#..#...#.#.#..#...#.#.#..#...#.#.#..#...#.#.#..#...#.#.#..#...#.#  --->
```

You start on the open square (.) in the top-left corner and need to reach the bottom (below the bottom-most row on your map).

The toboggan can only follow a few specific slopes (you opted for a cheaper model that prefers rational numbers); start by counting all the trees you would encounter for the slope right 3, down 1:

From your starting position at the top-left, check the position that is right 3 and down 1. Then, check the position that is right 3 and down 1 from there, and so on until you go past the bottom of the map.

The locations you'd check in the above example are marked here with O where there was an open square and X where there was a tree:

```
..##.........##.........##.........##.........##.........##.......  --->
#..O#...#..#...#...#..#...#...#..#...#...#..#...#...#..#...#...#..
.#....X..#..#....#..#..#....#..#..#....#..#..#....#..#..#....#..#.
..#.#...#O#..#.#...#.#..#.#...#.#..#.#...#.#..#.#...#.#..#.#...#.#
.#...##..#..X...##..#..#...##..#..#...##..#..#...##..#..#...##..#.
..#.##.......#.X#.......#.##.......#.##.......#.##.......#.##.....  --->
.#.#.#....#.#.#.#.O..#.#.#.#....#.#.#.#....#.#.#.#....#.#.#.#....#
.#........#.#........X.#........#.#........#.#........#.#........#
#.##...#...#.##...#...#.X#...#...#.##...#...#.##...#...#.##...#...
#...##....##...##....##...#X....##...##....##...##....##...##....#
.#..#...#.#.#..#...#.#.#..#...X.#.#..#...#.#.#..#...#.#.#..#...#.#  --->
```

In this example, traversing the map using this slope would cause you to encounter 7 trees.

Starting at the top-left corner of your map and following a slope of right 3 and down 1, how many trees would you encounter?

```{r}

# read the files as vector of string
input <- read.csv("./day03_input.txt", header = F)[,1]

# create a "char" matrix of "." and "#"
char_map <- strsplit(input, "") %>% 
  unlist() %>% 
  matrix(nrow = length(input), byrow = T)

# convert in to 0 and 1 (1 for the trees) 
int_map <- 1*(char_map == "#")

# lets subset the matrix with the trajectory

# the hill trajectory: 3 to the left and 1 to the bottom
shifts <- c(
  0,1,
  0,2,
  0,3,
  1,3
) %>% matrix(nrow=4, byrow=T)

# starting from top left
base_pos <- c(1,1)
trajectory <- base_pos

# until the last row
for(i in 1:(nrow(int_map)-1)){
  steps <- rep(base_pos,4) %>% 
    matrix(nrow = 4, byrow = T) +
    shifts

  base_pos <- steps[4,]

  trajectory <- rbind(trajectory, base_pos)
}

# the columns repeat adjusting
y_coords <- trajectory[,2] %% ncol(int_map)
y_coords[y_coords==0] <- ncol(int_map)

trajectory[,2] <- y_coords

sum(int_map[trajectory])

```

### Part 2

Time to check the rest of the slopes - you need to minimize the probability of a sudden arboreal stop, after all.

Determine the number of trees you would encounter if, for each of the following slopes, you start at the top-left corner and traverse the map all the way to the bottom:

Right 1, down 1.
Right 3, down 1. (This is the slope you already checked.)
Right 5, down 1.
Right 7, down 1.
Right 1, down 2.
In the above example, these slopes would find 2, 7, 3, 4, and 2 tree(s) respectively; multiplied together, these produce the answer 336.

What do you get if you multiply together the number of trees encountered on each of the listed slopes?

```{r}
# count trees in a trajectory
# receives a int matrix where "1" is a tree and "0" not
# receives a "shift pattern" for each step down
countTreesByShift <- function(shift.pattern, area.map){

  # starting from top left
  base_pos <- c(1,1)
  trajectory <- base_pos

  # apply the pattern until the bottom
  for(i in 1:(nrow(area.map)-1)){
    step <- base_pos + shift.pattern
    base_pos <- step
    trajectory <- rbind(trajectory, step)
  }

  # the columns repeat adjusting the bound
  y_coords <- trajectory[,2] %% ncol(area.map)
  y_coords[y_coords==0] <- ncol(area.map)
  trajectory[,2] <- y_coords
  # avoid out of index in the number of rows
  trajectory <- trajectory[trajectory[,1]<=nrow(area.map),]

  # subset the map with trajectory coordinates and sum
  return(sum(int_map[trajectory]))
}

# read the files as vector of string
input <- read.csv("./day03_input.txt", header = F)[,1]

# create a "char" matrix of "." and "#"
char_map <- strsplit(input, "") %>% 
  unlist() %>% 
  matrix(nrow = length(input), byrow = T)

# convert in to 0 and 1 (1 for the trees) 
int_map <- 1*(char_map == "#")

# test the patterns

# part 1
countTreesByShift(c(1,3),int_map)

# part 2
c(1,1,1,3,1,5,1,7,2,1) %>% 
  matrix(ncol=2, byrow = T) %>%
  split(1:nrow(.)) %>% 
  purrr::map_dbl(countTreesByShift, area.map=int_map) %>% 
  prod()

```
## Day 4: Passport Processing

### Part One

You arrive at the airport only to realize that you grabbed your North Pole Credentials instead of your passport. While these documents are extremely similar, North Pole Credentials aren't issued by a country and therefore aren't actually valid documentation for travel in most of the world.

It seems like you're not the only one having problems, though; a very long line has formed for the automatic passport scanners, and the delay could upset your travel itinerary.

Due to some questionable network security, you realize you might be able to solve both of these problems at the same time.

The automatic passport scanners are slow because they're having trouble detecting which passports have all required fields. The expected fields are as follows:

byr (Birth Year)
iyr (Issue Year)
eyr (Expiration Year)
hgt (Height)
hcl (Hair Color)
ecl (Eye Color)
pid (Passport ID)
cid (Country ID)

Passport data is validated in batch files (your puzzle input). Each passport is represented as a sequence of key:value pairs separated by spaces or newlines. Passports are separated by blank lines.

Here is an example batch file containing four passports:

ecl:gry pid:860033327 eyr:2020 hcl:#fffffd
byr:1937 iyr:2017 cid:147 hgt:183cm

iyr:2013 ecl:amb cid:350 eyr:2023 pid:028048884
hcl:#cfa07d byr:1929

hcl:#ae17e1 iyr:2013
eyr:2024
ecl:brn pid:760753108 byr:1931
hgt:179cm

hcl:#cfa07d eyr:2025 pid:166559648
iyr:2011 ecl:brn hgt:59in

The first passport is valid - all eight fields are present. The second passport is invalid - it is missing hgt (the Height field).

The third passport is interesting; the only missing field is cid, so it looks like data from North Pole Credentials, not a passport at all! Surely, nobody would mind if you made the system temporarily ignore missing cid fields. Treat this "passport" as valid.

The fourth passport is missing two fields, cid and byr. Missing cid is fine, but missing any other field is not, so this passport is invalid.

According to the above rules, your improved system would report 2 valid passports.

Count the number of valid passports - those that have all required fields. Treat cid as optional. In your batch file, how many passports are valid?

```{r}
library(tidyverse)

# read data as text (vector of string by lines)
input <- read_lines("./day04_input.txt")

# change a empty line to "|" (trick to split later)
input[input==""] <- "|"

# colapse the data and split the data 
# into a list of fields x value
passports <- input %>% 
  # colapse into one string
  str_c(collapse = " ") %>% 
  # split each passport data
  str_split("\\|") %>% 
  unlist() %>% 
  # for each passport split the fields
  str_split(" ") %>% 
  # pre-process the fields (spliting )
  map(function(line){
    line[line!=""] %>% 
      str_split(":", simplify = T) %>% 
      return()
  })

is.valid.passdata <- function(pass.data){
  fields <- pass.data[,1]
  # has all fields
  if (length(fields) == 8) return(T)
  # has all fields but cid
  if (length(fields)==7 & length(fields[fields=="cid"])==0 ) return(T)
  return(F)
}

passports %>% 
  map_lgl(is.valid.passdata) %>% 
  sum()

```

### Part two

The line is moving more quickly now, but you overhear airport security talking about how passports with invalid data are getting through. Better add some data validation, quick!

You can continue to ignore the cid field, but each other field has strict rules about what values are valid for automatic validation:

```
byr (Birth Year) - four digits; at least 1920 and at most 2002.
iyr (Issue Year) - four digits; at least 2010 and at most 2020.
eyr (Expiration Year) - four digits; at least 2020 and at most 2030.
hgt (Height) - a number followed by either cm or in:
 - If cm, the number must be at least 150 and at most 193.
 - If in, the number must be at least 59 and at most 76.
hcl (Hair Color) - a # followed by exactly six characters 0-9 or a-f.
ecl (Eye Color) - exactly one of: amb blu brn gry grn hzl oth.
pid (Passport ID) - a nine-digit number, including leading zeroes.
cid (Country ID) - ignored, missing or not.
```

Your job is to count the passports where all required fields are both present and valid according to the above rules. Here are some example values:

```
byr valid:   2002
byr invalid: 2003

hgt valid:   60in
hgt valid:   190cm
hgt invalid: 190in
hgt invalid: 190

hcl valid:   #123abc
hcl invalid: #123abz
hcl invalid: 123abc

ecl valid:   brn
ecl invalid: wat

pid valid:   000000001
pid invalid: 0123456789
```

Here are some invalid passports:

```
eyr:1972 cid:100
hcl:#18171d ecl:amb hgt:170 pid:186cm iyr:2018 byr:1926

iyr:2019
hcl:#602927 eyr:1967 hgt:170cm
ecl:grn pid:012533040 byr:1946

hcl:dab227 iyr:2012
ecl:brn hgt:182cm pid:021572410 eyr:2020 byr:1992 cid:277

hgt:59cm ecl:zzz
eyr:2038 hcl:74454a iyr:2023
pid:3556412378 byr:2007
```

Here are some valid passports:

pid:087499704 hgt:74in ecl:grn iyr:2012 eyr:2030 byr:1980
hcl:#623a2f

eyr:2029 ecl:blu cid:129 byr:1989
iyr:2014 pid:896056539 hcl:#a97842 hgt:165cm

hcl:#888785
hgt:164cm byr:2001 iyr:2015 cid:88
pid:545766238 ecl:hzl
eyr:2022

iyr:2010 hgt:158cm hcl:#b6652a ecl:blu byr:1944 eyr:2021 pid:093154719

Count the number of valid passports - those that have all required fields and valid values. Continue to treat cid as optional. In your batch file, how many passports are valid?

```{r}
library(tidyverse)

# read data as text (vector of string by lines)
input <- read_lines("day04_input.txt")

# change a empty line to "|" (trick to split later)
input[input==""] <- "|"

# colapse the data and split the data 
# into a list of fields x value
passports <- input %>% 
  # colapse into one string
  str_c(collapse = " ") %>% 
  # split each passport data
  str_split("\\|") %>% 
  unlist() %>% 
  # for each passport split the fields
  str_split(" ") %>% 
  # pre-process the fields
  map(function(line){
    # remove empty strings and split into "name":"value
    pd <- str_split(line[line!=""], ":", simplify = T)
    # return a list of fields
    pd[,2] %>% 
      split(1:length(.)) %>% 
      set_names(pd[,1]) %>% 
      as_tibble() %>% 
      return()
  })

# puts in a tibble
pass.check <- tibble(
    pdata = passports
  ) %>% 
  # unnest
  unnest(pdata) %>% 
  # validates each field
  # byr (Birth Year) - four digits; at least 1920 and at most 2002.
  mutate( byr = as.integer(byr) ) %>% 
  mutate( chk.byr = (!is.na(byr) & byr>=1920 & byr<=2002) ) %>% 
  # iyr (Issue Year) - four digits; at least 2010 and at most 2020.
  mutate( iyr = as.integer(iyr) ) %>% 
  mutate( chk.iyr = (!is.na(iyr) & iyr>=2010 & iyr<=2020) ) %>% 
  # eyr (Expiration Year) - four digits; at least 2020 and at most 2030.
  mutate( eyr = as.integer(eyr) ) %>% 
  mutate( chk.eyr = (!is.na(eyr) & eyr>=2020 & eyr<=2030)) %>% 
  # hgt (Height) - a number followed by either cm or in:
  mutate( hgt.fmt.chk = (!is.na(hgt) & str_detect(hgt,"[0-9]+(in|cm)")) ) %>% 
  # separates value of unit to check the ranges
  mutate( hgt.val = as.integer(str_extract(hgt,"[0-9]+")),
          hgt.unit = str_extract(hgt,"(in|cm)")) %>% 
  mutate( chk.hgt = case_when(
    # invalid format or empty field
    !hgt.fmt.chk ~ F,
    # - If cm, the number must be at least 150 and at most 193.
    hgt.unit == "cm" ~ (hgt.val>=150 & hgt.val<=193),
    # - If in, the number must be at least 59 and at most 76.
    hgt.unit == "in" ~ (hgt.val>=59 & hgt.val<=76),
    T ~ F # otherwise?
  )) %>% 
  # hcl (Hair Color) - a # followed by exactly six characters 0-9 or a-f.
  mutate( chk.hcl = (!is.na(hcl) & str_detect(hcl, "#[0-9a-f]{6}") & nchar(hcl)==7) ) %>% 
  # ecl (Eye Color) - exactly one of: amb blu brn gry grn hzl oth.
  mutate( chk.ecl = (!is.na(ecl) & ecl %in% c("amb","blu","brn","gry","grn","hzl","oth"))) %>% 
  # pid (Passport ID) - a nine-digit number, including leading zeroes.
  mutate( chk.pid = (!is.na(pid) & str_detect(pid,"[0-9]{9}") * nchar(pid)==9) ) %>%
  # cid (Country ID) - ignored, missing or not. 
  # check all policies
  mutate( is.valid = (chk.byr & chk.iyr & chk.eyr & chk.hgt & chk.hcl & chk.ecl & chk.pid))

# how many is valid?
pass.check %>% 
  count(is.valid)

```